---
title: "Advanced Regression"
author: "Walter McConnell"
date: "25 July 2018"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(reshape2)
library(plyr)
library(gridExtra)
library(grid)
```

```{r, echo=FALSE}
htmltools::img(src = "https://upload.wikimedia.org/wikipedia/commons/6/62/Intercom_logo.png", 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width: 200px;')
```


# Research Question
To predict the sales price for each house. For each Id in the test set, you must predict the value of the SalePrice variable.

### Research Evaluation Metric
The predictive model is evaluated on Root-Mean-Squared-Error (RMSE) between the logarithm of the predicted value and the logarithm of the observed sales price. 

### What is the scientific / business application ? 
To predict future house prices from collected data 


# Raw Data

```{r, include= FALSE, warning=FALSE}
test_data <- read_csv("data/raw/test.csv")
```

The data set describing the sale of individual residential property in Ames, Iowa from 2006 to 2010. The data set contains 2930 observations and a large number of explanatory variables (23 nominal, 23 ordinal, 14 discrete, and 20 continuous) involved in assessing home values. 

### How was the data collected ?
Collected by the Ames City Assessor’s Office as part of an 'assessment model' project. [Ames, Iowa: Alternative to the Boston Housing Data as an End of Semester Regression Project](https://ww2.amstat.org/publications/jse/v19n3/decock.pdf) 

### Raw Data Structure

#### Dimensions

```{r, echo = FALSE}
dim(test_data)
```

# Variables

```{r variables_table, echo = FALSE}

create_variables_table <- function(data) {
  types <- test_data %>% summarise_all(class) %>% gather(Name, class)
  unique_values <- test_data %>% summarise_all(function(x) length(unique(x))) %>% gather(Name, `Unique Values`)
  missing_count   <- test_data %>% summarise_all(function(x) sum(is.na(x))) %>% gather(Name, `Missing Data Count`)
  missing_percent <- test_data %>% summarise_all(function(x) round(mean(is.na(x)) * 100, 1)) %>% gather(Name, `Missing Data %`)

  join_all(list(types, unique_values, missing_count, missing_percent), by='Name', type='left')  
}

create_variables_table(test_data) %>% arrange(Name) %>% kable() 
```

## Numeric Variables

```{r, numeric_variables, echo = FALSE}

numeric_variables <- function() {
  create_variables_table(test_data) %>% filter(class %in% c("integer", "double", "numeric"))
}

numeric_variables() %>% .$Name %>% sort()
```

### Missing Data

```{r, echo = FALSE}
numeric_variables() %>% 
  arrange(desc(`Missing Data %`)) %>% 
  filter(`Missing Data %` > 0) %>% 
  kable()
```

#### LotFrontage

```{r, echo = FALSE}
numeric_variables() %>% filter(Name == "LotFrontage") %>% kable()
```

LotFrontage represents the `Linear feet of street connected to property`. LotFrontage has 15.6%/227 rows of missing data. 15.6% of rows is too much to drop, so will impute a value instead.

```{r, echo = FALSE, warning=FALSE}
histogram <- ggplot(test_data, aes(LotFrontage)) + 
  geom_histogram(bins = 50) +
  geom_freqpoly(bins = 40, color = "steelblue")

boxplot <- ggplot(test_data, aes(1, LotFrontage)) + 
  geom_boxplot() + 
  geom_point(position = "jitter", alpha = 0.5, color = "steelblue") +
  geom_hline(yintercept = mean(test_data$LotFrontage, na.rm = TRUE), color = 'pink')

grid.arrange(histogram, boxplot, ncol=2)
```

The median seems like a reasonable value to impute.

```{r, echo = FALSE}
LotFrontage_median <- median(test_data$LotFrontage, na.rm = TRUE)
test_data <- test_data %>% mutate(LotFrontage = ifelse(is.na(LotFrontage), LotFrontage_median, LotFrontage))

histogram <- ggplot(test_data, aes(LotFrontage)) + 
  geom_histogram(bins = 50) +
  geom_freqpoly(bins = 40, color = "steelblue")

boxplot <- ggplot(test_data, aes(1, LotFrontage)) + 
  geom_boxplot() + 
  geom_point(position = "jitter", alpha = 0.5, color = "steelblue") +
  geom_hline(yintercept = mean(test_data$LotFrontage, na.rm = TRUE), color = 'pink')

grid.arrange(histogram, boxplot, ncol=2)
```


#### GarageYrBlt

```{r, echo = FALSE}
numeric_variables() %>% filter(Name == "GarageYrBlt") %>% kable()
```

GarageYrBlt has 5.3%/58 missing values. It is in fact a categorical variable and so will be recoded as a factor. NA values will be reset to the same as the YearBuilt variable. Also there is an unreasonable outlier (a year of > 2200). This will also be adjusted.

```{r, echo = FALSE, warning = FALSE}
ggplot(test_data, aes(GarageYrBlt)) + geom_bar(fill = "steelblue") 
```


```{r, echo = FALSE}
test_data <- test_data %>% 
  mutate(GarageYrBlt = ifelse((is.na(GarageYrBlt) | GarageYrBlt > 2200), YearBuilt, GarageYrBlt))

test_data$GarageYrBlt <- factor(test_data$GarageYrBlt)
test_data$YearBuilt <- factor(test_data$YearBuilt)

ggplot(test_data, aes(GarageYrBlt)) + geom_bar(fill = "steelblue") 
```


#### MasVnrArea

```{r, echo = FALSE}
numeric_variables() %>% filter(Name == "MasVnrArea") %>% kable()
```

MasVnrArea refers to the "Masonry veneer area in square feet".

```{r, echo = FALSE, warning = FALSE}
histogram <- ggplot(test_data, aes(MasVnrArea)) + 
  geom_histogram(bins = 50) +
  geom_freqpoly(bins = 40, color = "steelblue")

boxplot <- ggplot(test_data, aes(1, MasVnrArea)) + 
  geom_boxplot() + 
  geom_point(position = "jitter", alpha = 0.5, color = "steelblue")

grid.arrange(histogram, boxplot, ncol=2)
```

When the MasVnrType variable is NA, the MasVnrArea variable is also na. 

MasVnrArea will be set to 0, and MasVnrType will be encoded as a factor, and set to 'None' for NA's.

```{r, echo = FALSE, warning = FALSE}
test_data <- test_data %>% 
  mutate(MasVnrArea = ifelse(is.na(MasVnrArea), 0 , MasVnrArea),
        MasVnrType = ifelse(is.na(MasVnrType), 'None', MasVnrType))

test_data$MasVnrType <- factor(test_data$MasVnrType)

histogram <- ggplot(test_data, aes(MasVnrArea)) + 
  geom_histogram(bins = 50) +
  geom_freqpoly(bins = 40, color = "steelblue")

boxplot <- ggplot(test_data, aes(1, MasVnrArea)) + 
  geom_boxplot() + 
  geom_point(position = "jitter", alpha = 0.5, color = "steelblue")

grid.arrange(histogram, boxplot, ncol=2)
```

#### Other NA numeric variables

```{r, echo = FALSE}
numeric_variables() %>% filter(Name %in% c("BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF", "TotalBsmtSF", "BsmtFullBath", "BsmtHalfBath")) %>% kable()
```

##### Basement related NA values

NA values are replaced with 0

```{r, echo= FALSE}
test_data <- test_data %>% mutate_at(vars(BsmtFinSF1, BsmtFinSF2, BsmtUnfSF, TotalBsmtSF, BsmtFullBath, BsmtHalfBath), function(x) ifelse(is.na(x),0,x))
```

#### Garage related NA values

```{r, echo = FALSE}
numeric_variables() %>% filter(Name %in% c("GarageCars", "GarageArea")) %>% kable()
```

NA values are replaced with 0

```{r, echo= FALSE}
test_data <- test_data %>% mutate_at(vars(GarageCars,GarageArea), function(x) ifelse(is.na(x),0,x))
```

All numeric variables have complete data 

```{r, echo = FALSE}
numeric_variables() %>% 
  arrange(desc(`Missing Data %`)) %>% 
  kable()
```


## Non-Numeric Variables

```{r, non_numeric_variables, echo = FALSE}

non_numeric_variables <- function() {
  create_variables_table(test_data) %>% filter(!class %in% c("integer", "double", "numeric"))
}

non_numeric_variables() %>% .$Name %>% sort()
```


### Non Numeric Variables - Variables with Missing Data

```{r, echo = FALSE}
non_numeric_variables() %>% 
  arrange(desc(`Missing Data %`)) %>% 
  filter(`Missing Data %` > 0) %>%
  kable()
```


#### Categorical variables with NA encoded as None

From the data description, some variables are encoded as NA to represent the absense of the feature (e.g PoolQC NA == 'no Pool'). 

BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2 = "No Basement"

FireplaceQu = "No Fireplace"
GarageType, GarageFinish, GarageQual, GarageCond,  = "No garage"
PoolQC = "No Pool"
Fence = "No Fence"
MiscFeature = "None"



```{r, echo = FALSE, include = FALSE}
test_data <- test_data %>% mutate(
  Alley = ifelse(is.na(Alley), "No alley access", Alley),
  BsmtQual = ifelse(is.na(BsmtQual), "No Basement", BsmtQual), 
  BsmtCond= ifelse(is.na(BsmtCond), "No Basement", BsmtCond), 
  BsmtExposure = ifelse(is.na(BsmtExposure), "No Basement", BsmtExposure), 
  BsmtFinType1 = ifelse(is.na(BsmtFinType1), "No Basement", BsmtFinType1), 
  BsmtFinType2 = ifelse(is.na(BsmtFinType2), "No Basement", BsmtFinType2),
  FireplaceQu = ifelse(is.na(FireplaceQu), "No Fireplace", FireplaceQu),
  GarageType = ifelse(is.na(GarageType), "No Garage", GarageType), 
  GarageFinish = ifelse(is.na(GarageFinish), "No Garage", GarageFinish), 
  GarageQual = ifelse(is.na(GarageQual), "No Garage", GarageQual), 
  GarageCond = ifelse(is.na(GarageCond), "No Garage", GarageCond),
  PoolQC = ifelse(is.na(PoolQC), "No Pool", PoolQC),
  Fence = ifelse(is.na(Fence), "No Fence",Fence),
  MiscFeature = ifelse(is.na(MiscFeature), "None", MiscFeature))

test_data$Alley <- factor(test_data$Alley)
test_data$BsmtQual <- factor(test_data$BsmtQual)
test_data$BsmtCond <- factor(test_data$BsmtCond)
test_data$BsmtExposure <- factor(test_data$BsmtExposure)
test_data$BsmtFinType1 <- factor(test_data$BsmtFinType1)
test_data$BsmtFinType2 <- factor(test_data$BsmtFinType2)
test_data$FireplaceQu <- factor(test_data$FireplaceQu)
test_data$GarageType <- factor(test_data$GarageType)
test_data$GarageFinish <- factor(test_data$GarageFinish)
test_data$GarageQual <- factor(test_data$GarageQual)
test_data$GarageCond <- factor(test_data$GarageCond)
test_data$PoolQC <- factor(test_data$PoolQC)
test_data$Fence <- factor(test_data$Fence)
test_data$MiscFeature <- factor(test_data$MiscFeature)
```

### Non Numeric Variables - Variables with Missing Data

```{r, echo = FALSE}
non_numeric_variables() %>% 
  arrange(desc(`Missing Data %`)) %>% 
  filter(`Missing Data %` > 0) %>%
  kable()
```


```{r, echo = FALSE}
test_data <- test_data %>% mutate(
  MSZoning = coalesce(MSZoning, names(sort(-table(test_data$MSZoning)))[1]),
  Utilities = coalesce(Utilities, names(sort(-table(test_data$Utilities)))[1]),
  Exterior1st = coalesce(Exterior1st, names(sort(-table(test_data$Exterior1st)))[1]),
  Exterior2nd = coalesce(Exterior2nd, names(sort(-table(test_data$Exterior2nd)))[1]),
  KitchenQual = coalesce(KitchenQual, names(sort(-table(test_data$KitchenQual)))[1]),
  Functional = coalesce(Functional, names(sort(-table(test_data$Functional)))[1]),
  SaleType = coalesce(SaleType, names(sort(-table(test_data$SaleType)))[1])
)
```

### Non Numeric Variables - Variables with Missing Data

```{r, echo = FALSE}
non_numeric_variables() %>% 
  arrange(desc(`Missing Data %`)) %>% 
  kable()
```
